Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-05-09 VegOzone.tex
Refine section 2.2.

\subsection{Ozone data}
\label{subsec:ozone_data}
To determine the significance of the higher \chem{[O_3]} in 2018 in relation to regional averages and trends, we derive a ground-level \chem{[O_3]} climatology for Svanvik (Section~\ref{sec:stats}). From 1986--1996, ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted before the 2018/19 growing season.

@@

As indicated above, all ground-level ozone observations used in this study are available from the EBAS database operated by \citet{NILU_EBAS}. A comprehensive overview over all past and present ozone observation sites in northern Fennoscandia is given in Table~\ref{tab:ebas_obs} and Fig.~\ref{fig:station_map_fennoscandia}, respectively. All stations with available long-term observations are located at higher altitude than Svanhovd which will result in relatively higher ground-level \chem{[O_3]} in comparison. In the absence of local precursors, this is due to the stratification of the atmosphere leading to an increase in ozone abundance with altitude \citep[e.g.,][]{AB:Klingberg2009}. The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath. However, at higher spatial resolution the vegetation varies substantially. Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above treeline (pine forest) and the surrounding vegetation mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}. At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass adhering to the \citet{WMOGuide2018} standards (Fig.\ref{fig:svanhovd_research_station}).

>>>>

\subsection{Ozone data}
\label{subsec:ozone_data}
From 1986--1996, ozone monitoring has been operated continuously by NILU at Svanhovd, but no ozone measurements have been conducted before the 2018/19 growing season. To determine the significance of the \chem{[O_3]} in 2018/19 in relation to regional trends, we derive a ground-level \chem{[O_3]} climatology for Svanvik and northern Fennoscandia in general (Section~\ref{sec:stats}).

@@

A comprehensive overview over all past and present ozone observation sites in northern Fennoscandia is given in Table~\ref{tab:ebas_obs} and Fig.~\ref{fig:station_map_fennoscandia}, respectively. All stations with available long-term observations are located at higher altitude than Svanhovd. In the absence of local precursors, the stratification of the atmosphere usually leads to a higher abundances in ozone at higher altitude \citep[e.g.,][]{AB:Klingberg2009}. The periphery vegetation at all sites is similar and consists mainly of pine forests, birch, and heath. However, on closer inspection the vegetation differs substantially. Ozone measurements at Pallas are taken on top of mount Sammaltunturi $100\,\unit{m}$ above treeline (pine forest) and the surrounding vegetation mainly consists of low vascular plants, mosses, and lichen \citep{BER:Hatakka2003}. At Svanhovd, the ozone measurements are taken amidst a meadow surrounded by agriculturally managed herbaceous plants and grass adhering to the \citet{WMOGuide2018} standards (Fig.\ref{fig:svanhovd_research_station}). In conjunction with the location at lower altitude, this results in generally lower \chem{[O_3]} at Svanhovd compared to Pallas.
